Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-04-15 Ben_Manuscripts/stochastic_transport/Final_Draft_JPCB.tex
Modifications based on MRS comments

  Up through our previous studies, our molecular models have provided valuable
  qualitative mechanistic insight with some quantitative support. This insight
  already allows us to speculate about new LLC monomer designs. However, they
  would be of greater value to a larger set of researchers if we could provide
  quantitative predictions of macroscopic observables such as solute flux and
  selectivity. Due to the size of the types of systems we are studying (at
  least 62,000 atoms) it is prohibitively expensive and time-consuming to run
  simulations longer than those performed for this work ($\sim$5 $\mu$s).  Even
  with the relatively large size of our system, we only simulate 24 solute
  trajectories per unit cell in order to minimize solute-solute interactions.
  This results in relatively high uncertainties in observables such as mean
  squared displacement (MSD), preventing reliable long timescale predictions.
  We are in need of a way of studying the collective motion of a much larger
  set of solutes over much longer timescales.
  %MRS11: Maybe this last sentence could be changed. Because you aren't really finding a way to study larger sets of 
  %solutes over larger timescales, you are using a limited amount of atomistic simulation into long time scale behavior efficiently. We can discuss variations as needed!

  %BJC: maybe this paragraph can be cut down?
  Mathematical descriptions of transport in complex separations membranes are a
  powerful way to understand mechanisms and formulate design principles.
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  %MRS11: this next sentence could probably be cut - it's kind of vague.
  The complexity of a well-fit model generally parallels the complexity of the
  transport mechanism being studied, as well as the transport information the
  model conveys. 
  In dense homogeneous membranes, the solution-diffusion model
  can extract diffusion and partition coefficients and has successfully
  predicted solute transport rates.~\cite{wijmans_solution-diffusion_1995}
  Analogously, pore-flow models yield predictions of diffusion coefficients and
  solute transport rates in nanoporous membranes.~\cite{paul_diffusive_1974}
  Modern single particle tracking approaches have taken researchers beyond
  continuum modeling allowing them to characterize complex diffusive
  behavior.~\cite{manzo_review_2015} At the molecular level, one can use
  molecular dynamics (MD) simulations to study both single particle dynamics
  and bulk transport properties with atomic-level
  insight.~\cite{coscia_chemically_2019,maginn_best_2018} All of these
  approaches facilitate generation of hypotheses about the molecular origins of
  separations by attempting to give a more intuitive understanding of how
  solutes move as a function of their environment, in turn suggesting
  experiments that could be performed.

@@

  Using a bottom-up modeling approach, we can parameterize single particle
  behavior by extracting particle trajectories from MD simulations and studying
  the properties that have the greatest influence on solute dynamics. With this
  information, we can learn how to construct an ensemble of characteristic
  single solute trajectories with these properties which would be useful for
  making long timescale predictions with computational ease and lower
  uncertainties. Fortunately, there is an abundance of information encoded in
  the complex solute trajectories. We can incorporate fluctuations and 
  time-dependent correlations in solute position into transport models, 
  as well as the length of time solutes are trapped. We can add further detail by integrating
  this time series analysis with our knowledge of the primary trapping
  mechanisms as well as the solute's changing chemical environment within the
  heterogeneous membrane.

@@

  %MRS11: should be clear if first model is previous publications, or what is presented here.
  %If what is presented here, should talk about how it will be presented, not that it was developed. 
  %MRS11: research activities happened in the past, presentation of research is described in the present.
  %or potentially future if describing what will be presented in the rest of the paper. 
  %BJC: replacement to above.
  We constructed our first model in order to characterize the hopping and
  trapping behavior displayed by solutes in our MD simulations. This type 
  of non-Brownian dynamical behavior results in subdiffusive transport which
  is well-described by anomalous diffusion (AD) theory.~\cite{metzler_random_2000,bouchaud_anomalous_1990,klages_anomalous_2008}

  % BJC: focused following paragraph (RWF doesn't need to be talked about) and 
  % stole thesis from paragraph below it
  We chose to treat our system in terms of two well-known classes of anomalous
  subdiffusion: fractional Brownian motion (FBM) subordinate to a 
  continuous time random walk (CTRW), or subordinated FBM (sFBM) for short. 
  FBM is common in crowded, viscoelastic environments where each jump comes 
  from a Gaussian distribution but is anti-correlated to its previous steps.
  ~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
  A pure CTRW is characterized by a distribution of hop lengths and dwell times, 
  where trajectories consist of sequential independent random draws from each
  distribution.~\cite{montroll_random_1965} If the draws are not independent,
  then the hop sequence generation process can be described as subordinate to
  the CTRW. In our case, we observe anti-correlated draws from the hop distribution
  suggesting an FBM process is subordinate to the CTRW.

@@

%  We treated our system as an FBM process subordinate to a CTRW, or
%  subordinated FBM (sFBM) for short. 
  We fit sFBM models in two ways: First, we parameterized solute motion using
  a single set of parameters fit to a hypothesized anomalous diffusion (AD) model.
  Second, we used a two-state approach by extracting two sets of model parameters
  calculated based on a solute's radial distance from the closest pore center. 
  This allows us to include the different dynamical behavior of solutes previously
  observed within the pore region versus within the monomer tails.

@@

  Our second modeling approach applies an extended MSM framework to a
  relatively small set of known states based on the three previously observed
  solute trapping mechanisms. Standard MSMs are typically applied to determine
  equilibrium populations of states and the kinetics of transitions between
  those states.~\cite{bowman_using_2009} We extend the framework to include
  state-dependent fluctuations and correlations in solute position. The
  magnitude of a solute's fluctuations from its average trapped position and the
  degree of correlation with previous fluctuations is determined by the current
  state. To distinguish our approach from standard MSMs, we have named it the
  Markov state-dependent dynamical model (MSDDM).

@@

  %MRS11: Introduction should not really give the answer so far.
  %Introduction should describe the questions that will bbe asked, but
  %not answer them.  That's not what is expected in an intro, it may confuse people (answers should be in the abstract).
  %MRS11: this next paragraph could use tweaks/cutting the most, but review the previous paragraphs as well.
  %MRS11: you can think about putting some of these results into conclusion, though - it reads well now, but might be benefitted from reemphasizing results if it's currently missing any of this.

  Finally, we use both models in order to predict solute flux and selectivity
  in pores of macroscopic length, thus achieving a better understanding of
  macroscopic properties on the basis of microscopic dynamics. We show that
  anti-correlated solute hopping behavior severely reduces flux relative to
  uncorrelated behavior. We use the relative solute fluxes to estimate the
  membrane's selectivity between all pairs of solutes and demonstrate that,
  when solutes display different degrees of anti-correlated hopping behavior,
  selectivity becomes a function of pore length. With this improved
  understanding of macroscopic behavior, we can begin to think more critically
  about how to design membranes in order to selectively pass or reject specific
  solutes.

>>>>

  Up through our previous studies, our molecular models have provided valuable
  qualitative mechanistic insight with some quantitative support. This insight
  already allows us to speculate about new LLC monomer designs. However, they
  would be of greater value to a larger set of researchers if we could provide
  quantitative predictions of macroscopic observables such as solute flux and
  selectivity. Due to the size of the types of systems we are studying (at
  least 62,000 atoms) it is prohibitively expensive and time-consuming to run
  simulations longer than those performed for this work ($\sim$5 $\mu$s). Even
  with the relatively large size of our system, we only simulate 24 solute
  trajectories per unit cell in order to minimize solute-solute interactions.
% BJC2: removed this sentence. High uncertainty MSDs don't really relate to long timescale predictions. Unless the MSDs were linear.
%  This results in relatively high uncertainties in observables such as mean
%  squared displacement (MSD), preventing reliable long timescale predictions.
  We are in need of an efficient method which can use our limited atomistic 
  simulation data sets in order to predict long timescale solute behavior.  
  %BJC2: following sentence replaced with above
%  We are in need of a way of studying the collective motion of a much larger
%  set of solutes over much longer timescales.
  %MRS11: Maybe this last sentence could be changed. Because you aren't really finding a way to study larger sets of 
  %solutes over larger timescales, you are using a limited amount of atomistic simulation into long time scale behavior efficiently. We can discuss variations as needed!

  %BJC: maybe this paragraph can be cut down?
  Mathematical descriptions of transport in complex separations membranes are a
  powerful way to understand mechanisms and formulate design principles.
  \cite{vinh-thang_predictive_2013,geens_transport_2006,darvishmanesh_mass_2016}
  %MRS11: this next sentence could probably be cut - it's kind of vague.
  %BJC2: True
%  The complexity of a well-fit model generally parallels the complexity of the
%  transport mechanism being studied, as well as the transport information the
%  model conveys. 
  In dense homogeneous membranes, the solution-diffusion model
  can extract diffusion and partition coefficients and has successfully
  predicted solute transport rates.~\cite{wijmans_solution-diffusion_1995}
  Analogously, pore-flow models yield predictions of diffusion coefficients and
  solute transport rates in nanoporous membranes.~\cite{paul_diffusive_1974}
  Modern single particle tracking approaches have taken researchers beyond
  continuum modeling allowing them to characterize complex diffusive
  behavior.~\cite{manzo_review_2015} At the molecular level, one can use
  molecular dynamics (MD) simulations to study both single particle dynamics
  and bulk transport properties with atomic-level
  insight.~\cite{coscia_chemically_2019,maginn_best_2018} All of these
  approaches facilitate generation of hypotheses about the molecular origins of
  separations by attempting to give a more intuitive understanding of how
  solutes move as a function of their environment, in turn suggesting
  experiments that could be performed.

@@

  Using a bottom-up modeling approach, it may be possible to parameterize single
  particle behavior by extracting particle trajectories from MD simulations and
  studying the properties that have the greatest influence on solute dynamics.
  With this information, one can learn how to construct an ensemble of characteristic
  single solute trajectories with these properties which would be useful for
  making long timescale predictions with computational ease and lower
  uncertainties. Fortunately, there is an abundance of information encoded in
  the complex solute trajectories. One can incorporate fluctuations and 
  time-dependent correlations in solute position into transport models, 
  as well as the length of time solutes are trapped. It is possible to add further
  detail by integrating this time series analysis with existing knowledge of the
  primary trapping mechanisms as well as the solute's changing chemical environment
  within the heterogeneous membrane.

@@

  %MRS11: should be clear if first model is previous publications, or what is presented here.
  %If what is presented here, should talk about how it will be presented, not that it was developed. 
  %MRS11: research activities happened in the past, presentation of research is described in the present.
  %or potentially future if describing what will be presented in the rest of the paper. 
  %BJC: replacement to above.
  %BJC1: Made clear that this is talking about the first model of this work.
  %BJC1: went through rest of intro to make sure tense is consistent
  The first model that we develop in this work, allows us to characterize the 
  hopping and trapping behavior displayed by solutes in our MD simulations. This 
  type of non-Brownian dynamical behavior results in subdiffusive transport which
  is well-described by anomalous diffusion (AD) theory.~\cite{metzler_random_2000,bouchaud_anomalous_1990,klages_anomalous_2008}

  % BJC: focused following paragraph (RWF doesn't need to be talked about) and 
  % stole thesis from paragraph below it
  We treat our system in terms of two well-known classes of anomalous
  subdiffusion: fractional Brownian motion (FBM) subordinate to a 
  continuous time random walk (CTRW), or subordinated FBM (sFBM) for short. 
  FBM is common in crowded, viscoelastic environments where each jump comes 
  from a Gaussian distribution but is anti-correlated to its previous steps.
  ~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
  A pure CTRW is characterized by a distribution of hop lengths and dwell times, 
  where trajectories consist of sequential independent random draws from each
  distribution.~\cite{montroll_random_1965} If the draws are not independent,
  then the hop sequence generation process can be described as subordinate to
  the CTRW. In our case, we observe anti-correlated draws from the hop distribution
  suggesting an FBM process is subordinate to the CTRW.

@@

%  We treated our system as an FBM process subordinate to a CTRW, or
%  subordinated FBM (sFBM) for short. 
  We fit sFBM models in two ways: First, we parameterize solute motion using
  a single set of parameters fit to a hypothesized anomalous diffusion (AD) model.
  Second, we use a two-state approach by extracting two sets of model parameters
  calculated based on a solute's radial distance from the closest pore center. 
  This allows us to include the different dynamical behavior of solutes previously
  observed within the pore region versus within the monomer tails.

@@

  Our second modeling approach applies an extended MSM framework to a
  relatively small set of known states based on the three previously observed
  solute trapping mechanisms. Standard MSMs are typically applied to determine
  equilibrium populations of states and the kinetics of transitions between
  those states.~\cite{bowman_using_2009} We extend the framework to include
  state-dependent fluctuations and correlations in solute position. The model
  determines the magnitude of a solute's fluctuations from its average trapped
  position and the degree of correlation with previous fluctuations by its current
  state. To distinguish our approach from standard MSMs, we have named it the
  Markov state-dependent dynamical model (MSDDM).

@@

  %MRS11: Introduction should not really give the answer so far.
  %Introduction should describe the questions that will bbe asked, but
  %not answer them.  That's not what is expected in an intro, it may confuse people (answers should be in the abstract).
  %MRS11: this next paragraph could use tweaks/cutting the most, but review the previous paragraphs as well.
  %MRS11: you can think about putting some of these results into conclusion, though - it reads well now, but might be benefitted from reemphasizing results if it's currently missing any of this.
  %BJC2: slight tweaks and removed conclusions from following paragraph. Also went back through previous
  % paragraphs to look for conclusions.

  Finally, we demonstrate how one can use stochastically generated realizations
  of our models in order to predict solute flux and selectivity in pores of 
  macroscopic length, thus achieving a better understanding of macroscopic 
  properties on the basis of microscopic dynamics. By improving our 
  understanding of macroscopic behavior in this way, we can begin to think more
  critically about how to design membranes in order to selectively pass or 
  reject specific solutes.
